In this project we are trying to solve the automated musical composition problem using Deep Reinforcement Learning. Solving this problem with the usage of Reinforcement Learning becomes an interesting and innovative approach since it enables the interaction between the computer and humans for teaching which compositions sound good and that way improve the computer's compositional skills. This can become a difficult task to automate, since there is no specific way to measure if a composition is better or worse than another one. Music quality depends on each user's preferences, because of that, a composition that may sound nice for a group of people, may not sounds as nice as a different composition according to a different group of people's opinion. Based on this, we consider that it is important to create a bridge between computational musical composition and human feedback, reason why Reinforcement Learning can give promising results on this approach.

In this report, we will present the results that we have obtained by applying \emph{Q-learning} and \emph{Deep Q-learning}. We were able to teach the computer which notes are the best to be played after other notes according to the user's criteria by asking them for their feedback after a note or sequence of notes have been played. That way, we were able to generate a melody that match the user's preferences. We will give information about past works regarding musical encoding for computational composition and the usage of reinforcement learning in this area. After that, we will provide information about the approach used for solving this problem and explain the results obtained. Finally, we will mention future steps that we could take in this project.